Estimating migration rates for iteroparous species using kinship pairs found between populations

### 2021-12-14 latex/ms.tex
fix Eqs.13-14
add range of M in main text

Suppose there are $N_1$ parents in the population 1 at the beginning of the first year. Each parent produces offspring in which the number is governed by the parent's reproductive potential, denoted by $\lambda_{i,1}$ ($i=1,2,\ldots,N_1$). After the reproductive season, part of the parents would start to moving towards the population 2; $M$ survived migrants arrive in the population 2 so that there are $N_2$ parents in the population 2 at the beginning of the second year ($M < N_2$). Similar to the population 1, $N_2$ parents produce offspring in which the number is governed by $\lambda_{j,2}$ ($j=1,2,\ldots,N_2$). The reproductive potential is determined by several factors and its details are summarized in {\bf Appendix}; theoretically speaking, the number of offspring for each parent is set as a random variable with mean $\lambda_{i,1}$ or $\lambda_{j,2}$.

@@

To assess these probabilities, recall there must be two parents of arbitral offspring; the first probability is the sum of father-offspring and mother-offspring relationship, which can be written by $r_{\rm M}M/(r_{\rm S,1}N_{1}) + (1 - r_{\rm M})M/((1-r_{\rm S,1})N_{1})$, where $r_{\rm M}$ and $r_{\rm S,1}$ indicates sex ratio of parents in the migrants and in the whole of the population 1, respectively. Assuming $r_{\rm M} = r_{\rm S,1}$, the probability is simplified as $2M/N_{1}$. Similarly, the second probability is simplified as $2M/N_{2}$, implying the assumption of equal sex ratio in the migrants and the others. The third probability corresponds to the probability that randomly selected two migrants are identical, that is $1/M$. Taken together, we obtain
\begin{align}
\pi_{\rm HS,bet} &= \frac{2M}{N_{1}} \frac{2M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{4M}{N_{1}N_{2}}.
\label{HS-1}
\end{align}
This form also holds under a flexible setting for the reproductive potential, such as a situation with random variable of $\lambda_{i,1}$ and $\lambda_{j,2}$, as noted in {\bf Appendix}.

@@

Next, we consider the probability that offspring sampled in the population 1 and a parent sampled in the population 2 share an PO relationship, denoted by $\pi_{\rm PO, bet}$. $\pi_{\rm PO, bet}$ can be partitioned into the three probabilities: (i) the probability that sampled offspring in the population 1 born to a migrant; (ii) the probability that a migrant is sampled; and (iii) the probability that the migrant of a sampled offspring in the population 1 and the migrant sampled in the population 2 are identical. The first and second probabilities are the same as the one introduced in the previous subsection, that is, $2M/N_{1}$ and $1/M$. The second probability is $M/N_{2}$ by definition. Taken together, we obtain
\begin{align}
\pi_{\rm PO,bet} &= \frac{2M}{N_{1}} \frac{M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{2M}{N_{1}N_{2}}.
\label{PO-1}
\end{align}

@@

When the both HS pairs and PO pairs, which are found between populations, are simultaneously available, we can obtain efficient estimators by combined $\widehat{M_1}$ and $\widehat{M_2}$ for migration number: 
\begin{align}
\widehat{M_3} &= \frac{ N_{1} N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right), }
\label{AP-1}
\end{align}
and by combining $\widehat{m_1}$ and $\widehat{m_2}$ for migration rate: 
\begin{align}
\widehat{m_3} &= \frac{ N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right). }
\label{AP-2}
\end{align}
For those estimators, $\widetilde{H}_{\rm HS,bet}$ and $\widetilde{H}_{\rm PO,bet}$ is weighted by a sample size $n_{\rm O, 2}$ and $n_{\rm P, 2}$, respectively.

>>>>

Suppose there are $N_1$ parents in the population 1 at the beginning of the first year. Each parent produces offspring in which the number is governed by the parent's reproductive potential, denoted by $\lambda_{i,1}$ ($i=1,2,\ldots,N_1$). After the reproductive season, part of the parents would start to moving towards the population 2; $M$ survived migrants arrive in the population 2 so that there are $N_2$ parents in the population 2 at the beginning of the second year ($M \le N_1$ and $M \le N_2$). Similar to the population 1, $N_2$ parents produce offspring in which the number is governed by $\lambda_{j,2}$ ($j=1,2,\ldots,N_2$). The reproductive potential is determined by several factors and its details are summarized in {\bf Appendix}; theoretically speaking, the number of offspring for each parent is set as a random variable with mean $\lambda_{i,1}$ or $\lambda_{j,2}$.

@@

To assess these probabilities, recall there must be two parents of arbitral offspring; the first probability is the sum of father-offspring and mother-offspring relationship, which can be written by $r_{\rm M}M/(r_{\rm S,1}N_{1}) + (1 - r_{\rm M})M/((1-r_{\rm S,1})N_{1})$, where $r_{\rm M}$ and $r_{\rm S,1}$ indicates sex ratio of parents in the migrants and in the whole of the population 1, respectively. Assuming $r_{\rm M} = r_{\rm S,1}$, the probability is simplified as $2M/N_{1}$. Similarly, the second probability is simplified as $2M/N_{2}$, implying the assumption of equal sex ratio in the migrants and the others. The third probability corresponds to the probability that randomly selected two migrants are identical, that is $1/M$. Taken together, we obtain
\begin{align}
\pi_{\rm HS,bet} &= \frac{2M}{N_{1}} \frac{2M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{4M}{N_{1}N_{2}}.
\label{HS-1}
\end{align}
This form holds under a flexible setting for the reproductive potential, such as a situation with random variable of $\lambda_{i,1}$ and $\lambda_{j,2}$, as noted in {\bf Appendix}. If all $N_1$ parents safely move to the population 2 (i.e., $M = N_1$), $\pi_{\rm HS,bet}$ equals $4/N_2$; meanwhile, if the population 2 consists only of migrants (i.e., $M = N_2$), $\pi_{\rm HS,bet}$ equals $4/N_1$. These probabilities in the extreme cases are similar forms to HS probability that is randomly sampled from two different cohorts within a population \cite[]{bravington2016close}.

@@

Next, we consider the probability that offspring sampled in the population 1 and a parent sampled in the population 2 share an PO relationship, denoted by $\pi_{\rm PO, bet}$. $\pi_{\rm PO, bet}$ can be partitioned into the three probabilities: (i) the probability that sampled offspring in the population 1 born to a migrant; (ii) the probability that a migrant is sampled; and (iii) the probability that the migrant of a sampled offspring in the population 1 and the migrant sampled in the population 2 are identical. The first and second probabilities are the same as the one introduced in the previous subsection, that is, $2M/N_{1}$ and $1/M$. The second probability is $M/N_{2}$ by definition. Taken together, we obtain
\begin{align}
\pi_{\rm PO,bet} &= \frac{2M}{N_{1}} \frac{M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{2M}{N_{1}N_{2}}.
\label{PO-1}
\end{align}
This form also holds under a flexible setting for the reproductive potential, as noted in {\bf Appendix}. If all $N_1$ parents safely move to the population 2 (i.e., $M = N_1$), $\pi_{\rm PO,bet}$ equals $2/N_2$; meanwhile, if the population 2 consists only of migrants (i.e., $M = N_2$), $\pi_{\rm PO,bet}$ equals $2/N_1$. These probabilities in the extreme cases are similar forms to PO probability that is randomly sampled within a population \cite[]{bravington2016close}.

@@

When the both HS pairs and PO pairs, which are found between populations, are simultaneously available, we can obtain efficient estimators by combined $\widehat{M_1}$ and $\widehat{M_2}$ for migration number: 
\begin{align}
\widehat{M_3} &= \frac{ N_{1} N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) },
\label{AP-1}
\end{align}
and by combining $\widehat{m_1}$ and $\widehat{m_2}$ for migration rate: 
\begin{align}
\widehat{m_3} &= \frac{ N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right)}.
\label{AP-2}
\end{align}
For those estimators, $\widetilde{H}_{\rm HS,bet}$ and $\widetilde{H}_{\rm PO,bet}$ is weighted by a sample size $n_{\rm O, 2}$ and $n_{\rm P, 2}$, respectively.
